What can we learn about life in the city by mining the Melbourne’s Open Data Platform?

### 2017-05-07 thesis/proposal_draft.Rmd
set alpha = 1 in plot legend for figure 6

The pedestrian data is presented as a time series. However, we face a major obstacle with pedestrian data: multiple seasonal patterns. Looking at the data on a month to month basis at a single sensor location, we have two main cycles: a daily sub-cycle within a weekly cycle.
```{r, message= FALSE, warning= FALSE, fig.height=2, echo = FALSE, fig.align= "center"}
bsplot <- ped_df %>% filter(Year == "2016", Month == "June", Mdate < 25, Sensor_ID == "13")
ggplot(bsplot) + geom_line(aes(x = MTime, y = Hourly_Counts)) +
      xlab("Hour of the Month") + ylab("Hourly Counts") + theme_light()
```
\begin{center}
\textbf{Figure 5}:  Sample of hourly counts of pedestrians at Flagstaff Station (June 2016). Two repeating cycles can be seen, with a public holiday on 13 July showing atypical pedestrian counts for a weekday (Monday).
\end{center}  
From _Figure 5_, we can observe multiple characteristics of the time series. Firstly, we see the weekly cycle, where weekdays (Monday to Friday) have lots of activity and weekends (Saturday and Sunday) are relatively quiet. This would be highly suggestive that Flagstaff Station is primarily used by city office workers. Another phenomenon we observe is the effect of a public holiday. Looking near the middle of the month (Monday, 13 June 2016), we have the Queen's Birthday public holiday. We must also note that, prior to the start of 2016, Flagstaff Station was not operating on weekends, nor public holidays.
```{r, message= FALSE, warning= FALSE, fig.height=3, echo = FALSE, fig.align= "center"}
ggplot(filter(ped_df, Sensor_ID == "13")) + geom_path(aes(x = as.integer(Time), y = Hourly_Counts, colour = Day), alpha = 0.2) +
      xlab("Hour of the Day") + ylab("Hourly Counts") + theme_light()

```
\begin{center}
 \textbf{Figure 6}:  Hourly counts at Flagstaff Station, with colour mapped to day of the week - There are two distinctly different daily patterns. The typical weekday traffic pattern has two peaks at 9am and 5pm, while it remains relatively quiet over the weekend as seen by the flatter lines. Also note that the second peak is lower than the first peak on average.
\end{center}

>>>>

The pedestrian data is presented as a time series. However, we face a major obstacle with pedestrian data: multiple seasonal patterns. Looking at the data on a month to month basis at a single sensor location, we have two main cycles: a daily sub-cycle within a weekly cycle.
```{r, message= FALSE, warning= FALSE, fig.height=2, echo = FALSE, fig.align= "center"}
bsplot <- ped_df %>% filter(Year == "2016", Month == "June", Mdate < 25, Sensor_ID == "13")
ggplot(bsplot) + geom_line(aes(x = MTime, y = Hourly_Counts)) +
      xlab("Hour of the Month") + ylab("Hourly Counts") + theme_light()
```
\begin{center}
\textbf{Figure 5}:  Sample of hourly counts of pedestrians at Flagstaff Station (June 2016). Two repeating cycles can be seen, with a public holiday on 13 July showing atypical pedestrian counts for a weekday (Monday).
\end{center}  
From _Figure 5_, we can observe multiple characteristics of the time series. Firstly, we see the weekly cycle, where weekdays (Monday to Friday) have lots of activity and weekends (Saturday and Sunday) are relatively quiet. This would be highly suggestive that Flagstaff Station is primarily used by city office workers. Another phenomenon we observe is the effect of a public holiday. Looking near the middle of the month (Monday, 13 June 2016), we have the Queen's Birthday public holiday. We must also note that, prior to the start of 2016, Flagstaff Station was not operating on weekends, nor public holidays.
```{r, message= FALSE, warning= FALSE, fig.height=3, echo = FALSE, fig.align= "center"}
ggplot(filter(ped_df, Sensor_ID == "13")) + geom_path(aes(x = as.integer(Time), y = Hourly_Counts, colour = Day), alpha = 0.2) +
      xlab("Hour of the Day") + ylab("Hourly Counts") + theme_light() + guides(colour = guide_legend(override.aes = list(alpha = 1)))
```
\begin{center}
 \textbf{Figure 6}:  Hourly counts at Flagstaff Station, with colour mapped to day of the week - There are two distinctly different daily patterns. The typical weekday traffic pattern has two peaks at 9am and 5pm, while it remains relatively quiet over the weekend as seen by the flatter lines. Also note that the second peak is lower than the first peak on average.
\end{center}
